---
title: "The Modeling Process"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
# Set global knitr chunk options
knitr::opts_chunk$set(
  fig.align = "center",
  error = FALSE,
  message = FALSE, 
  warning = FALSE, 
  collapse = TRUE 
)
```

-   the machine learning process is very iterative and heurstic-based

-   common for many ML approaches to be applied, evaluated, and modified before a final, optimal model can be determined

-   A proper process needs to be implemented to have confidence in our results

## Components of the process

Concepts that are useful for any type of machine learning model:

-   data splitting
-   fitting models
-   resampling
-   bias-variance trade-off
-   making predictions
-   model evaluation

## Prerequisites

Packages

```{r prereq-pkg}
library(tidyverse)
library(tidymodels)
```

Data

Assuming you've downloaded the class data and you have it in a subdirectory titled "data".

```{r prereq-data}
library(here)
data_path <- here('Module 8/ML-data-1')

# ames data
ames <- read_csv(here(data_path, "ames.csv"))

# attrition data
attrition <- read_csv(here(data_path, "attrition.csv")) %>%
  mutate(Attrition = factor(Attrition))
```

## Regression problem

```{r}
ames %>%
  select(Sale_Price, everything())
```

## Classification problem

```{r}
attrition %>%
  select(Attrition, everything())
```

## Data Splitting

**Generalizability**: we want an algorithm that not only fits well to our past data, but more importantly, one that predicts a future outcome accurately.

-   Training Set: these data are used to develop feature sets, train our algorithms, tune hyper-parameters, compare across models, and all of the other activities required to reach a final model decision.

-   Test Set: having chosen a final model, these data are used to estimate an unbiased assessment of the model’s performance (generalization error).

## Mechanics of data splitting

```{r split-ames}
set.seed(123)
ames_split <- initial_split(ames, prop = 0.7, strata = Sale_Price)
ames_train <- training(ames_split)
ames_test  <- testing(ames_split)
```

```{r split-attrit}
set.seed(123)
churn_split <- initial_split(attrition, prop = 0.7, strata = Attrition)
churn_train <- training(churn_split)
churn_test  <- testing(churn_split)
```

## Validating split distribution

We should always validate that our train and test splits have similar distributions:

```{r, fig.height=3}
# Do the distributions line up? 
ggplot(ames_train, aes(x = Sale_Price)) + 
  geom_line(stat = "density", 
            trim = TRUE) + 
  geom_line(data = ames_test, 
            stat = "density", 
            trim = TRUE, col = "red")
```

```{r}
# consistent response ratio between train & test
table(churn_train$Attrition) %>% prop.table()

table(churn_test$Attrition) %>% prop.table()
```

## Building models

We can use the Tidymodels approach, which is a meta engine (aggregator) that allows you to apply almost any direct engine

```{r}
# Tidymodels meta engine
lm_tidy <- linear_reg() %>%
  fit(Sale_Price ~ Gr_Liv_Area + Year_Built, data = ames_train)
```

### 3 Steps:

1.  Create a model object

2.  Choose an "engine"

3.  Fit the model

### Step 1: create a model object

```{r}
my_model <- linear_reg() 
```

-   `linear_reg()`: Linear regression

-   `logistic_reg()`: Logistic regression

-   `decision_tree()`: Decision tree

-   `rand_forest()`: Random forest

-   `nearest_neighbor()`: K-nearest neighbor

-   `naive_Bayes()`: Naive Bayes

-   and [many more](https://parsnip.tidymodels.org/reference/index.html#models)!

### Step 2: choose an engine

```{r}
lm_model <- linear_reg() %>%
  set_engine('lm') 

glm_model <- linear_reg() %>%
  set_engine('glm') 

keras_model <- linear_reg() %>%
  set_engine('keras') 
```

-   The "engine" is just the underlying package we want to use to train our model.

-   Every model object has a default engine (i.e. `linear_reg()` defaults to use **lm**).

-   But we can change the engine used if desired.

### Step 3: Fit the model

For our regression model...

```{r}
lm_model <- linear_reg() %>%
  set_engine('lm') %>%
  fit(Sale_Price ~ Gr_Liv_Area + Year_Built, data = ames_train) 
```

And for our classification model...

```{r}
classifier_model <- logistic_reg() %>%
  fit(Attrition ~ ., data = churn_train) 
```

## Making predictions

So we call this ***training*** or ***fitting*** our model:

```{r}
lm_model <- linear_reg() %>%
  set_engine('lm') %>%
  fit(Sale_Price ~ Gr_Liv_Area + Year_Built, data = ames_train)
```

Once we have a model, we can use it to make predictions whether that be on...

the current training data:

```{r}
lm_model %>% predict(ames_train)
```

Or on new, unseen data:

```{r}
lm_model %>% predict(ames_test)
```

## Model evaluation

**Regression**

-   Mean Square Error (MSE)
-   Root Mean Square Error (RMSE)
-   Mean Absolute Error (MAE)
-   Mean Absolute Percent Error (MAPE)
-   Root Mean Squared Logarithmic Error (RMSLE)

```{r}
lm_model %>%
  predict(ames_test) %>%
  bind_cols(ames_test %>% select(Sale_Price)) %>%
  rmse(truth = Sale_Price, estimate = .pred)
```

**Classification**

-   Classification Accuracy
-   Recall vs. Specificity
-   $F_1$ Score
-   Log Loss

```{r}
classifier_model %>%
  predict(churn_test) %>%
  bind_cols(churn_test %>% select(Attrition)) %>%
  accuracy(truth = Attrition, estimate = .pred_class)
```

```{r}
classifier_model %>%
  predict(churn_test) %>%
  bind_cols(churn_test %>% select(Attrition)) %>%
  conf_mat(truth = Attrition, estimate = .pred_class)
```

## Summary

1.  Split data
2.  Train model
3.  Make predictions
4.  Use predictions to evaluate performance

```{r}
# split data
set.seed(123)
ames_split <- initial_split(ames, prop = 0.7, strata = Sale_Price)
ames_train <- training(ames_split)
ames_test  <- testing(ames_split)

# train model
lm_model <- linear_reg() %>%
  set_engine('lm') %>%
  fit(Sale_Price ~ Gr_Liv_Area + Year_Built, data = ames_train)

# make predictions
predictions <- lm_model %>% 
  predict(ames_test)

# evaluate model performance
predictions %>% 
  bind_cols(ames_test %>% select(Sale_Price)) %>%
  rmse(truth = Sale_Price, estimate = .pred)
```
