---
title:
output:
  html_notebook: default
  html_document:
    theme: cerulean
    toc: true
    toc_depth: 2
---
# Report {.tabset}

## Introduction

The business problem we're addressing focuses on how different types of coupon campaigns affect sales across various product departments and display locations at Regork. Understanding which coupon types lead to higher sales is crucial for optimizing promotions, improving customer engagement, and boosting revenue.

We utilized sales transaction data, focusing on coupon types and their redemption rates, campaign display locations, and total sales across different departments. The methodology included analyzing sales distributions across these variables to identify patterns in consumer purchases.

This analysis offers actionable insights, such as which coupon types drive the most sales and which display locations are most effective. This can help Regork stakeholders make data-driven decisions on future promotions, improve resource allocation, and enhance overall sales strategies, especially by focusing on high-performing areas like grocery or drugstore departments.

## Packages Required
```{r}
library(tidyverse)
library(here)
library(lubridate)
library(ggplot2)
library(readr)
library(completejourney)
library(readr)
```
## Data Preparation {.tabset}

### Setup
```{r}
c(promotions, transactions) %<-% get_data(which = 'both', verbose = FALSE)
# setwd('/Users/ezishr/Documents/CINCY/Fall 2024/BANA 4080/BANA4080_midterm_project')

transactions <- transactions %>% mutate(date = as.Date(transaction_timestamp))
transactions$product_id <- as.double(transactions$product_id)
promotions$product_id <- as.double(promotions$product_id)
products$product_id <- as.double(products$product_id)
coupons$product_id <- as.double(coupons$product_id)

display_location_labels <- c("0"="No display", 
                             "1"="Store front", 
                             "2"="Store rear", 
                             "3"="Front end cap", 
                             "4"="Mid-aisle end cap", 
                             "5"="Rear end cap", 
                             "6"="Side aisle end cap", 
                             "7"="In-aisle", 
                             "9"="Secondary location display", 
                             "A"="In-shelf")

mailer_location_labels <- c("0"="Not on ad", 
                            "A"="Interior page feature", 
                            "C"="Interior page line item", 
                            "D"="Front page feature", 
                            "F"="Back page feature", 
                            "H"="Wrap front feature", 
                            "J"="Wrap interior coupon", 
                            "L"="Wrap back feature", 
                            "P"="Interior page coupon", 
                            "X"="Free on interior page", 
                            "Z"="Free on front page/back page/or wrap")

```

### With Redemption
Assumption: transactions with any discount are redeemed coupons
```{r}
trans1 <- transactions %>% 
  dplyr::filter(retail_disc>0 | coupon_disc>0 | coupon_match_disc>0) %>%
  select(household_id, store_id, product_id, sales_value:coupon_match_disc, week, date) 

# Join coupons, coupon_redemptions, campaign_description to get full info of coupons
redemptions <- coupons %>%
  inner_join(campaign_descriptions) %>%
  inner_join(coupon_redemptions, relationship = 'many-to-many') %>%
  rename(date = redemption_date)

# Every transaction with redemption
sample_df <- redemptions %>%
  left_join(trans1, by = c('household_id','product_id','date'), relationship = 'many-to-many')

colSums(is.na(sample_df)) # there are roughly 2M coupons redemption w/o transaction records

transWithRedemptions <- redemptions %>%
  inner_join(trans1, by = c('household_id','product_id','date'), relationship = 'many-to-many') %>%
  arrange(desc(product_id))

# Total sales from transactions with redemption, grouped by product_id and store_id
transWithRedemptions_groupedByProductStoreId <- transWithRedemptions %>%
  group_by(product_id, store_id) %>%
  summarise(total_sales = sum(sales_value))

transWithRedemptions_grouped <- transWithRedemptions_groupedByProductStoreId

#### Check if there is duplicate comb of product_id and store_id in every row
duplicates <- transWithRedemptions_grouped %>%
  dplyr::group_by(product_id, store_id) %>%
  dplyr::summarize(count = n()) %>%
  dplyr::filter(count > 1)
print(duplicates)


# Get unique existing product id and store id from the above
product_info_id <- unique(transWithRedemptions_groupedByProductStoreId$product_id)
store_info_id <- unique(transWithRedemptions_groupedByProductStoreId$store_id)

# Get rid of location 0 (not display) for display_location
promos_display <- promotions %>% 
  select(-(mailer_location:week)) %>% 
  dplyr::distinct(product_id, store_id, display_location, .keep_all = TRUE) %>%
  dplyr::filter(display_location != 0)

# Join display location to transactions with redemption by keys: product_id and store_id
transWithRedemptions_displayLocation <- transWithRedemptions_groupedByProductStoreId %>%
  inner_join(promos_display, by=c('product_id', 'store_id')) %>%
  left_join(products, by='product_id') %>%
  select(-c('manufacturer_id', 'brand','package_size'))

# Get rid of location 0 for mailer_location
promos_mailer <- promotions %>% 
  select(-c(display_location, week)) %>% 
  dplyr::distinct(product_id, store_id, mailer_location, .keep_all = TRUE) %>%
  dplyr::filter(mailer_location != 0)

# Join mailer location to transactions with redemption by keys: product_id and store_id
transWithRedemptions_mailerLocation <- promos_mailer %>%
  inner_join(transWithRedemptions_groupedByProductStoreId, by=c('product_id', 'store_id')) %>%
  left_join(products, by='product_id') %>%
  select(-c('manufacturer_id', 'brand','package_size'))

# Promotions with display_location = 0, which is not displayed
promos_display0 <- promotions %>% 
  select(-c(mailer_location, week)) %>% 
  dplyr::filter(display_location == 0) %>%
  dplyr::distinct(product_id, store_id, .keep_all = TRUE)
  
# Join display location = 0 to transactions with redemption by keys product_id and store_id
transWithRedemptions_displayLocationIs0 <- transWithRedemptions_grouped %>%
  semi_join(promos_display0, by = c("product_id", "store_id")) %>%
  left_join(products, by='product_id') %>%
  select(-c('manufacturer_id', 'brand','package_size'))

# Promotions with mailer_location = 0, which is not displayed
promos_mailer0 <- promotions %>% 
  select(-c(display_location, week)) %>% 
  dplyr::filter(mailer_location == 0) %>%
  dplyr::distinct(product_id, store_id, .keep_all = TRUE)

# Join mailer location to transactions with redemptions by keys product_id and store_id
transWithRedemptions_mailerLocationIs0 <- transWithRedemptions_grouped %>%
  semi_join(promos_mailer0, by = c("product_id", "store_id")) %>%
  left_join(products, by='product_id') %>%
  select(-c('manufacturer_id', 'brand','package_size'))
```

### Without Redemption
```{r}
# Every transaction without redemption
trans_without_redemption <- transactions %>%
  dplyr::filter(retail_disc==0 & coupon_disc==0 & coupon_match_disc==0)

# Total sales with grouped by keys: product_id and store_id
transWithoutRedemption_Grouped <- trans_without_redemption %>% group_by(product_id, store_id) %>% summarise(total_sales = sum(sales_value), .groups="drop")

### Check duplicates
transWithoutRedemption_Grouped %>% group_by(product_id, store_id) %>% summarise(count = n()) %>% dplyr::filter(count > 1)

# Add info about products to trans
transWithoutRedemption_Grouped <- transWithoutRedemption_Grouped %>% 
  left_join(products, by='product_id') %>%
  select(product_id, store_id, total_sales, department, product_category, product_type)
```

### Export as csv files
```{r}
full_transactions <- transWithRedemptions
trans_w_display <- transWithRedemptions_displayLocation
trans_w_mailer <- transWithRedemptions_mailerLocation
trans_w_display0 <- transWithRedemptions_displayLocationIs0
trans_w_mailer0 <- transWithRedemptions_mailerLocationIs0
```

## Exploratory Data Analysis {.tabset}
### Coupon Types Effect on Sales
- Assuming 'campaign_type' contains different coupon types

We first tried to understand the performances of each campaign type on its redemption counts and total sales value. We achieved this by first joining the coupons related datasets to get every information of coupons. Then, we joined it with the transactions dataset to filter those having coupons redeemed, named full_transactions.

We proceeded to group this by campaign_type in a bid to outline the trend of the totals sales value and redemption counts. This thus highlights Campaign Type A coming far ahead of others campaigns in both redemption count and total sales value. This implies how strong Campaign Type A was in triggering engagement from customers with the products having coupons.

With this in mind, further exposition is called for with regard to the inner mechanics of Campaign Type A. Which products it works best for, the structure of the campaign in and of itself, or what elements make it successful could all help inform actionable insights into how to better construct future campaigns. That could be more attractive discounts, better product placement, or even broader appeal across segments, all of which should be factored in for further analysis.

```{r, fig.width=13, fig.height=12}
summary_transactions <- full_transactions %>%
  group_by(campaign_type) %>%
  summarise(
    total_sales = sum(sales_value),
    count_type = n()
)

ggplot(data = summary_transactions, aes(x = campaign_type)) + 
  geom_col(aes(y = count_type, fill = 'Count of campaign type'), width = 0.5) +
  geom_line(aes(y = total_sales, color = 'Total Sales'), group = 1) +
  geom_point(aes(y = total_sales, color = 'Total Sales'), size = 3.0) + 
  scale_y_continuous(
    name = "Count of campaign type",
    breaks = seq(0, max(summary_transactions$total_sales) + 1, by = 400),
    sec.axis = sec_axis(~., 
                        name = "Total Sales",
                        breaks = seq(0, max(summary_transactions$total_sales) + 1, by = 400))
  ) +
  labs(x = 'Campaign Types', title = "Total Sales Value and Count of Redemptions Per Campaign Type") + 
  scale_fill_manual(name = '', values = c("Count of campaign type" = "lightblue")) +
  scale_color_manual(name = '', values = c("Total Sales" = 'red')) + 
  theme_minimal() + 
  theme(legend.position = 'bottom',
        axis.title.x = element_text(margin = margin(t = 10)),
        axis.title.y.left = element_text(margin = margin(r = 10)),
        axis.title.y.right = element_text(margin = margin(l = 10)),
        ) +
  annotate("text", x = "Type A", y = 7603, label = "Exceptionally High", color = "blue", size = 6, vjust = -1)
```

Thoughts:

- Type A has significantly higher redemption counts and sales value => might have more look into what type A is, how it works, which products that type A works on most.

```{r, fig.width=12}
products$product_id <- as.double(products$product_id)

sample <- full_transactions %>%
  group_by(product_id, campaign_type) %>%
  summarise(total = sum(sales_value), .groups="drop") %>%
  left_join(products, by="product_id") %>%
  select(campaign_type, department, total) %>%
  group_by(campaign_type, department) %>%
  summarise(total_new = sum(total), .groups="drop")
  
graph1 <- ggplot(data = sample, aes(x = campaign_type, y = department, fill = total_new)) +
  geom_tile() +
  scale_fill_gradient2(low="white", mid="lightblue", high="blue", name = "Sales Value") +
  geom_text(aes(label=round(total_new, 2)), color="black", size=4) +
  theme_minimal() +
  labs(title = "Distribution of Sales Value by Campaign Types and Departments",
       x = "Campaign Types",
       y = "Deparments") +
  theme(axis.title.x = element_text(margin = margin(t = 10)),
        axis.title.y = element_text(margin = margin(r = 10)),
)
```
- Thoughts:

  + The Grocery department stays on top of total sales value regardless of campaign types. This can be explained by being daily necessities. In the following EDA, we can omit Grocery department to see how other departments are doing.
  + Type A campaign has coupons redeemed mostly and in every department.

### Display Location Analysis
#### Hypothesis: Stuff at certain display locations with coupons will have the most purchase.
- As mentioned above, we exclude Grocery department in this visualization.
```{r, fig.width=15, fig.height=8}
sample <- trans_w_display %>%
  dplyr::filter(department != "GROCERY") %>%
  group_by(display_location, department) %>%
  summarize(total = sum(total_sales), .groups = 'drop')

print(sample, n = 5)
```

##### a. Stack bar plot with raw sales value for each display location
```{r, fig.width=15, fig.height=10}
ggplot(data = sample, aes(x = display_location, y = total, fill = department)) + 
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = c("DELI" = "#FF9999", 
                               "DRUG GM" = "#66CC99", 
                               "MEAT-PCKGD" = "#FFCC00", 
                               "NUTRITION" = "#3399FF", 
                               "MEAT" = "#FF66CC", 
                               "PASTRY" = "#FF9966", 
                               "SEAFOOD-PCKGD" = "#99CCFF")) +
  scale_x_discrete(labels = display_location_labels) +
  theme_minimal() +
  labs(title = "Total Sales by Department and Campaign Type in Percentage",
       x = "Display Locations",
       y = "Sales Value",
       fill = "Departments") +
  theme(axis.title.x = element_text(margin = margin(t = 10)),
        axis.title.y = element_text(margin = margin(r = 10)),
)
```

##### b. Stack bar plot with sales value being normalize to percentage
```{r, fig.width=15, fig.height=8}
total_sales_per_display_location <- sample %>%
  group_by(display_location) %>%
  summarise(total_sales_location = sum(total), .groups = "drop")

final_df <- sample %>%
  left_join(total_sales_per_display_location, by = "display_location") %>%
  mutate(percentage = round((total / total_sales_location) * 100,2))

ggplot(final_df, aes(x = display_location, y = percentage, fill = department)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = c("DELI" = "#FF9999", 
                               "DRUG GM" = "#66CC99", 
                               "MEAT-PCKGD" = "#FFCC00", 
                               "NUTRITION" = "#3399FF", 
                               "MEAT" = "#FF66CC", 
                               "PASTRY" = "#FF9966", 
                               "SEAFOOD-PCKGD" = "#99CCFF")) +
  scale_x_discrete(labels = display_location_labels) +
  theme_minimal() +
  labs(title = "Total Sales by Department and Campaign Type in Percentage",
       x = "Display Locations",
       y = "Percentage Sales",
       fill = "Departments") +
  theme(axis.title.x = element_text(margin = margin(t = 10)),
        axis.title.y = element_text(margin = margin(r = 10)),
)
```
- Thoughts:
  + Drug GM seems to have the most sales after Grocery department

#### Comparisons between being displayed and not displayed - Having coupons redeemed
##### Display
```{r, fig.width=15, fig.height=10}
sample1 <- trans_w_display %>%
  group_by(department) %>%
  summarize(total = sum(total_sales), .groups = 'drop') %>%
  arrange(desc(department))

ggplot(data = sample1, aes(x = department, y = total)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  coord_flip() + 
  scale_y_continuous(breaks = seq(0, max(sample1$total)+1, by=200)) +
  geom_text(aes(label = round(total, 2)),
            hjust = -0.1,
            color = "black") +
  labs(title = "Total Sales Across Department for Being Displayed",
       x = "Departments",
       y = "Total Sales") +
  theme_minimal() + 
  theme(axis.title.x = element_text(margin = margin(t = 10)))
```

##### Not Display
```{r, fig.width=15, fig.height=10}
sample2 <- trans_w_display0 %>%
  group_by(department) %>%
  summarise(total = sum(total_sales)) %>%
  arrange(desc(department))

ggplot(data = sample2, aes(x = department, y = total)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  coord_flip() + 
  scale_y_continuous(breaks = seq(0, max(sample2$total)+1, by=200)) +
  geom_text(aes(label = round(total, 2)),
            hjust = -0.1,
            color = "black") +
  labs(title = "Total Sales Across Department for Not Being Displayed",
       x = "Departments",
       y = "Total Sales") +
  theme_minimal() + 
  theme(axis.title.x = element_text(margin = margin(t = 10)))
```
- Thoughts:
  + Even without being displayed, Grocery department still has the highest sales value, following by Drug GM department. However, Meat Packaged is no longer at third position. This can implicit that people tends to buy packaged products if being displayed, otherwise people buy fresh meat.
  
### Mailer Location Analysis
#### Hypothesis: Customer is likely to redeem coupons if products are placed at interior page feature in mail.
```{r}
sample <- trans_w_mailer %>%
  group_by(mailer_location, department) %>%
  summarise(total = sum(total_sales), .groups = "drop") %>%
  arrange(desc(total))

print(sample, n=5)
```

##### a. Stack bar plot with raw sales value for each mailer location
```{r, fig.width=15, fig.height=12}
sample <- sample %>% dplyr::filter(department != "GROCERY")


ggplot(data = sample, aes(x = mailer_location, y = total, fill = department)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_x_discrete(labels = mailer_location_labels) + 
  scale_fill_manual(values = c("PRODUCE" = "#66c2a5",   
                               "GROCERY" = "#fc8d62",  
                               "DRUG GM" = "#8da0cb",      
                               "MEAT-PCKGD" = "#e78ac3",   
                               "MEAT" = "#a6d854",         
                               "NUTRITION" = "#ffd92f",     
                               "DELI" = "#e5c494",          
                               "SEAFOOD" = "#b3b3b3",      
                               "PASTRY" = "#1f78b4",       
                               "COSMETICS" = "#33a02c",    
                               "SEAFOOD-PCKGD" = "#fb9a99",
                               "TRAVEL & LEISURE" = "#e31a1c",
                               "COUPON" = "#ff7f00")) +
  scale_y_continuous(breaks = seq(0, 3500, by = 500), expand = c(0,2)) +
  theme_minimal() + 
  theme(axis.title.x = element_text(margin = margin(t=10)),
        axis.title.y = element_text(margin = margin(r=10))) +
  labs(title = "Total Sales by Department and Campaign Type",
       x = "Mailer Locations",
       y = "Total Sale",
       fill = "Departments")

```
Thoughts:

- TBD

##### b. Stack bar plot with sales value being normalize to percentage
```{r, fig.width=12, fig.height=8} 
total_sales_per_display_location <- sample %>%
  group_by(mailer_location) %>%
  summarise(total_sales_location = sum(total), .groups = "drop")

final_df <- sample %>%
  left_join(total_sales_per_display_location, by = "mailer_location") %>%
  mutate(percentage = round((total / total_sales_location) * 100,2))

ggplot(final_df, aes(x = mailer_location, y = percentage, fill = department)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = c("PRODUCE" = "#66c2a5",   
                               "GROCERY" = "#fc8d62",  
                               "DRUG GM" = "#8da0cb",      
                               "MEAT-PCKGD" = "#e78ac3",   
                               "MEAT" = "#a6d854",         
                               "NUTRITION" = "#ffd92f",     
                               "DELI" = "#e5c494",          
                               "SEAFOOD" = "#b3b3b3",      
                               "PASTRY" = "#1f78b4",       
                               "COSMETICS" = "#33a02c",    
                               "SEAFOOD-PCKGD" = "#fb9a99",
                               "TRAVEL & LEISURE" = "#e31a1c",
                               "COUPON" = "#ff7f00")) +
  scale_x_discrete(labels = mailer_location_labels) +
  theme_minimal() +
  scale_y_continuous(expand = c(0,2)) +
  theme(axis.title.x = element_text(margin = margin(t=10)),
        axis.title.y = element_text(margin = margin(r=10))) +
  labs(title = "Total Sales by Department and Campaign Type in Percentage",
       x = "Mailer Locations",
       y = "Percentage Sales",
       fill = "Departments")
```
Thoughts:

- TBD

#### Comparisons between being displayed and not displayed in mail - Having coupons redeemed
##### Mail Displayed
```{r, fig.width=15, fig.height=10}
sample1 <- trans_w_mailer %>%
  dplyr::filter(department!="GROCERY") %>%
  group_by(department) %>%
  summarize(total = sum(total_sales), .groups = 'drop') %>%
  arrange(desc(department))

ggplot(data = sample1, aes(x = department, y = total)) +
  geom_bar(stat = "identity", fill = "purple", width=0.5) +
  coord_flip() + 
  geom_text(aes(label = round(total, 2)),
            hjust = -0.1,
            color = "black") +
  labs(title = "Total Sales Across Department for Being Not Displayed in Mail",
       x = "Departments",
       y = "Total Sales") +
  theme_minimal() + 
  theme(axis.text.x = element_blank())
```
##### Not Mail Displayed
```{r, fig.width=15, fig.height=10}
sample2 <- trans_w_mailer0 %>%
  dplyr::filter(department != "GROCERY") %>%
  group_by(department) %>%
  summarise(total = sum(total_sales)) %>%
  arrange(desc(department))

ggplot(data = sample2, aes(x = department, y = total)) +
  geom_bar(stat = "identity", fill = "purple", width=0.5) +
  coord_flip() + 
  geom_text(aes(label = round(total, 2)),
            hjust = -0.1,
            color = "black") +
  labs(title = "Total Sales Across Department for Not Being Displayed in Mail",
       x = "Departments",
       y = "Total Sales") +
  theme_minimal() + 
  theme(axis.text.x = element_blank())
```

### Display vs. Mailer Promotions
```{r}
mailer_promo <- trans_w_mailer %>%
  select(department, total_sales) %>%
  group_by(department) %>%
  summarise(total = sum(total_sales), .groups="drop") %>%
  mutate(promo_type = "mail")

display_promo <- trans_w_display %>%
  ungroup() %>%
  select(department, total_sales) %>%
  group_by(department) %>%
  summarise(total = sum(total_sales), .groups="drop") %>%
  mutate(promo_type = "display")

summary_data <- bind_rows(mailer_promo, display_promo)
```

```{r, fig.width=15, fig.height=20}
ggplot(data = summary_data, aes(x = department, y = total, fill = promo_type)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("lightblue", "orange")) +
  geom_text(aes(label = round(total,2)), position = position_dodge(width = 0.9), vjust = -0.3, color = "black")
```

```{r, fig.width=13, fig.height=12}
ggplot(data = summary_data, aes(x = total, y = department, color = promo_type)) +
  geom_point(alpha = 0.6, size = 5) + 
  scale_x_log10() +
  labs(title = "Sales Comparison Between Display and Mailer Promotions",
       x = "Department",
       y = "Total Sales",
       color = "Exhibit Types") +
  theme_classic()
```








