---
title: "Writing Functions"
output: html_notebook
---

A function centralizes a common task to a single, abstract method

This helps to:

- increase reusability which...
- reduces verbosity
- reduces repetitiveness
- reduces the chance of making an error
- reduces instances requiring updates when changes are needed
- reduces amount of code to test

# Key ingredients to a function

1. Name
2. arguments
3. body
4. environment

```{r slide-7}
# present value function
pv <- function(fv, r, n = 5) {
  fv / (1 + r)^n
}

# arguments
args(pv)

# body
body(pv)

# environment
environment(pv)
```

## Your Turn!

Identify the arguments, body, and environment of:

1. `read.csv`
2. `tidyr::contains`
3. `sum`


# Define a function

- We define a function with <- just like we define any other R object
- Use informative names; strive to use verbs when possible
- We can define a function with no arguments; however, this is rarely useful
- Consequently, most functions have all key ingredients

```{r slide-13}
compute_pv <- function(fv, r, n) {
  fv / (1 + r)^n
}

compute_pv(fv = 1000, r = .05, n = 10)

compute_pv(fv = 2000, r = .08, n = 5)
```

## Your Turn!

- Define a function titled ratio that takes arguments x and y and returns their 
  ratio, `x / y`
- Call `ratio()` with arguments 3 and 4

```{r slide-14}
ratio <- function(___, ___) {
  _______
}
```


Calling arguments in different ways

Many ways to call arguments:

- Using argument names
- Positional matching
- Must use names if you change order
- ...otherwise error or incorrect computation will occur
- missing arguments results in error

```{r slide-15-18}
# using arg names
compute_pv(fv = 1000, r = .05, n = 10)

# positional matching
compute_pv(1000, .05, 10)

# use names when changing order
compute_pv(r = .05, fv = 1000, n = 10)

# ...otherwise an error occurs
compute_pv(.05, 1000, 10)

# missing arguments result in error
compute_pv(1000, .05)
```

# Setting good defaults

- As our functions require more inputs...
- many of them can likely do with defaults
- Usually, the first 1-3 arguments are not preset but the rest are
- Good defaults are largely based on
   - methodology best practices (i.e. hyperparameters, removing missing values)
   - industry best practices or current values (i.e. current interest rates)
   - organizational best practices (i.e. golden rules)

```{r slide-20}
# set default value for digits argument
compute_pv <- function(fv, r, n, digits = 2) {
  round(fv / (1 + r)^n, digits = digits)
}

# by default, will round to 2 digits
compute_pv(1000, .05, 10)

# to adjust, specify by position
compute_pv(1000, .05, 10, 0)

# or by name
compute_pv(1000, .05, 10, digits = 0)
```

## Your Turn Part 1

- Earlier in these slides you saw the following code duplicated:

```
(df$a - min(df$a, na.rm = TRUE)) / (max(df$a, na.rm = TRUE) - min(df$a, na.rm = TRUE))
```

- Can you write a function called rescale() that takes argument x and executes 
  this code?
- Test it on the following vector:

```{r slide-21}
set.seed(123)
test_vector <- runif(20, min = 25, max = 40)
```

## Your Turn Part 2

- Now add an argument to rescale() that allows you to round the output to a 
  specified decimal.
- Set the default to 2.


# Checking arguments and other conditions

- We've created a function that allows users to specify their inputs
- But how do we ensure they provide us with the right kind of inputs?
- Or what if we want to be able to provide some sort of feedback to the user?
- There are several ways to signal conditions to function users:
   - `stop()` & `stopifnot()`: signal an error (no way for a function to continue 
      and execution must stop)
   - `warning()`: signal a warning (something has gone wrong but the function has 
      been able to at least partially recover.)
   - `message()`: signal an informative message (function works fine but user 
      should be informed of something.)

## `stop()` & `stopifnot()`

- The stop functions are most commonly used to check for proper inputs
- but can be used to stop the function procedures because the user's environment 
  is not properly established
- Weigh the benefits of highly custom, informative error messages (via `stop`) 
  versus short to the point (via 
  `stopifnot`)
- Execution will stop at the first violation

```{r slide-24}
# stop
compute_pv <- function(fv, r, n, digits = 2) {
  if(!is.numeric(fv)) {
    stop("`fv` must be numeric", call. = FALSE)
  }
  if(!is.numeric(r)) {
    stop("`r` must be numeric", call. = FALSE)
  }
  if(!is.numeric(n)) {
    stop("`n` must be numeric", call. = FALSE)
  }
  if(!is.numeric(digits)) {
    stop("`digits` must be numeric", call. = FALSE)
  }
  round(fv / (1 + r)^n, digits = digits)
}

compute_pv("1000", ".05", 10)

# stopifnot
compute_pv <- function(fv, r, n, digits = 2) {
  stopifnot(is.numeric(fv), is.numeric(r), is.numeric(n), is.numeric(digits))
  round(fv / (1 + r)^n, digits = digits)
}
compute_pv("1000", ".05", 10)
```

## `warning()`

- `warnings()` are rarely used but...
- can be useful to signal non-terminating concerns
   - hyperparameter clashes in a grid search
   - an extremely long compute time and the user may want to re-assess
   - deprecated functions
   - NAs, NULLs, Infs exist, which result in NA output

```{r slide-25}
# example in established function
lubridate::mdy(c("2-1-2019", "1995"))

# how to implement
present_value <- function(fv, r, n, digits = 2) {
  
  warning("`present_value()` is deprecated. Use `compute_pv()`", call. = FALSE)
  stopifnot(is.numeric(fv), is.numeric(r), is.numeric(n), is.numeric(digits))
  
  round(fv / (1 + r)^n, digits = digits)
}

present_value(1000, .05, 10)
```

## `message()`

- designed to be informational
- use them to tell the user that you’ve done something on their behalf
- may represent:
  - default argument that may be non-trivial
  - how information was interpreted by the function
  - words of encouragement!
  - or even industry/business best practices
- Always provide an option to suppress messages

```{r slide-31}
compute_pv <- function(fv, r, n, digits = 2, quiet = FALSE) {
  
  stopifnot(is.numeric(fv), is.numeric(r), is.numeric(n), is.numeric(digits))
  
  if(!quiet) {
    if(r < 0.01 || r > 0.25) { 
    message("Your interest rate exceeds the normal ranges of 1-25%") 
    }
    # other messages here...
  }
  
  round(fv / (1 + r)^n, digits = digits)
}
# default provides message
compute_pv(1000, .26, 10)

# user can silence
compute_pv(1000, .26, 10, quiet = TRUE)
```

## Your Turn!

Let's go back to our `rescale()` function:

```{r slide-32a}
rescale <- function(x, digits = 2){
  rng <- range(x, na.rm = TRUE)
  scaled <- (x - rng[1]) / (rng[2] - rng[1])
  round(scaled, digits = digits)
}
```

- Move the `na.rm = TRUE` option to the function arguments so that the user can 
  control whether or not to remove missing values.
- Add a `warning()` that tells the user NAs are present (if they use `na.rm = FALSE`)
- Add a `message()` that tells the user how many NAs were removed
- add a missing value to your test_vector and test out your revised function


# Distributing your functions

Two main ways to distribute your functions:

1. using the source() function to read an R script
   - easy to get started
   - all functions/objects in the sourced .R script will be imported to your 
     global environment
   - does not scale well
   
```{r slide-35}
# you may need to adjust this path
fx_path <- here("my-functions", "custom-stat-functions.R")
source(fx_path)
ls()

purrr::map_dbl(mtcars, skewness)
```

2. writing a package (package writing will not be covered)
   - once you have a few functions that serve a common use
   - hard to get started
   - scales beautifully 
   

# Learning More

Advanced R: https://adv-r.hadley.nz/
